Multipoint Approximation Method for Large Scale Design Optimization Problems Under Uncertainty

### 2017-10-25 2018/LION/LION2018.tex
short version (without superquantile)

% EUROGEN 2015 - Template for the extended abstract
%defining format issues
%**********************

%Document Settings___________________________________________________________________

\documentclass[10pt,twocolumn,a4paper]{article}
\special{papersize=210mm,297mm}

%\makeatletter
%\renewcommand\section{\@startsection{section}{1}{\z@}%
%                                  {-3.5ex \@plus -1ex \@minus -.2ex}%
%                                  {2.3ex \@plus.2ex}%
%                                  {\normalfont\normalsize\bfseries}}
%\makeatother

%\makeatletter
%\renewcommand\subsection{\@startsection{subsection}{1}{\z@}%
%                                  {-3.5ex \@plus -1ex \@minus -.2ex}%
%                                  {2.3ex \@plus.2ex}%
%                                  {\normalfont\normalsize\em}}
%\makeatother

%________________________________________________________________________________

%Miscellanneous__________________________________________________________________

\sloppy % better line breaks
\newcommand{\authorname}{\textbf}
\newcommand{\authorgroup}{\emph}
\newcommand{\authoraddress}{\emph}
\newcommand{\authormail}{\emph}
%________________________________________________________________________________

%Packages________________________________________________________________________

%\usepackage[latin1]{inputenc}
\usepackage[utf8]{inputenx}
\usepackage[compact]{titlesec}
%\usepackage[pdftex]{graphicx}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,mathrsfs}
\usepackage{fancyhdr}
\usepackage{calc}
\usepackage{titling}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage[superscript]{cite}
\usepackage{titlesec}
\usepackage[none]{hyphenat}
\usepackage{enumitem}
\usepackage{cite}
%________________________________________________________________________________
\titleformat*{\section}{\normalfont\normalsize\bfseries}
\titleformat*{\subsection}{\normalfont\normalsize\em}
\voffset = -1cm
\textheight = 23.7cm
\hoffset = -0.5cm
\textwidth = 17cm
\setlength\parindent{0.5cm}
%Headers_________________________________________________________________________

\pagestyle{fancy}
\fancyhead[L]{EUROGEN 2017}\fancyhead[R]{September 13-15, 2017, Madrid, Spain}
\fancyheadoffset[LE,RO]{\marginparsep + \marginparwidth}
\setlength{\topmargin}{0pt}
%________________________________________________________________________________

%Maketitle Conf__________________________________________________________________
\pretitle{\begin{center}\Large}
\posttitle{\end{center}}

\preauthor{\begin{center}\normalsize \lineskip 0.5em\begin{tabular}[t]{c}}
\postauthor{\end{tabular}\par\end{center}}

\predate{\begin{center}\tiny}
\postdate{\par\end{center}}
%________________________________________________________________________________
%--------------------------------------------------------------------------------

%Article data entry_____________________________________________________________

\title{\vspace{-5ex}\textbf{Multipoint Approximation Method for Large Scale Design Optimization Problems Under Uncertainty}}
\author{
  \authorname{Vassili V. Toropov*}\\
  \authorgroup{Professor of Aerospace Engineering, School of Engineering and Materials Science}\\
  \authoraddress{Queen Mary University of London, London, UK}\\
  \authormail{Email: v.v.toropov@qmul.ac.uk}\\\\
  \authorgroup{Senior Research Fellow, Institute of Information Technology, Mathematics and Mechanics}\\
  \authoraddress{Lobachevsky University of Nizhny Novgorod, Russia}\\\\
  %--------------------------
  \authorname{Yury M. Korolev}\\
  \authorgroup{Research Fellow, Institute of Mathematics and Image Computing}\\
  \authoraddress{University of Lübeck, Germany}\\
  \authormail{Email: yury.korolev@mic.uni-luebeck.de}\\\\
  %--------------------------
  \authorname{Victor P. Gergel, Konstantin A. Barkalov}\\
  \authorgroup{Institute of Information Technology, Mathematics and Mechanics}\\
  \authoraddress{Lobachevsky University of Nizhny Novgorod, Russia}
  %}
}
\date{}

%_______________________________________________________________________________

%Article Body___________________________________________________________________

@@

%Abstract goes here--------------------------------------------------------------

\twocolumn[

@@

\thispagestyle{fancy} %Insert header in title page

\begin{center}
\line(1,0){465}
\end{center}

\begin{@twocolumnfalse}

\textbf{Summary}\\

TBD\\

\emph{\textbf{Keywords:} TBD}
\begin{center}
\line(1,0){465}
\end{center}

\end{@twocolumnfalse}

]

%--------------------------------------------------------------------------------

@@

Optimisation under uncertainty deals with responses $F(\pmb x,\pmb y)$ that depend on deterministic design variables $\pmb x$ and random ‘environmental’ variables $\pmb y$. Uncertainty in the design variables can also be modelled by introducing additional random variables, e.g. as additive or multiplicative noise. In this case, the responses $F(\pmb x,\pmb y)$ receive as input the sum $\pmb x+\pmb u$ (if the noise is additive), where $\pmb u$ is a random variable with a known probability distribution, yielding responses depending on $\pmb x,\; \pmb u$ and $\pmb y$. Randomness in $\pmb u$ and $\pmb y$ induces randomness in the responses $F(\pmb x+\pmb u,\pmb y)$. To convert them into deterministic functions, that can be optimised, a mapping $R$, called the \textit{risk measure}\cite{RockafellarUryasev2000} or \textit{robustness measure}\cite{SchillingsSchulz2015}, is applied to the responses. The result is then a deterministic function of the design variables: $\widetilde{F}(\pmb x)=R(F(\pmb x+\pmb u, \pmb y))$. The choice of a particular risk measure $R$ reflects the designer's attitude towards the risk of sub-optimal performance or constraint violation. For example, $R$ can be a combination of the mean and standard deviation of $F(\pmb x+ \pmb u,\pmb y)$ with respect to the joint probability distribution of $\pmb u$ and $\pmb y$ or a quantile of the distribution of $F(\pmb x+ \pmb u,\pmb y)$, i.e. such a value that the random variable of $F(\pmb x+\pmb u,\pmb y)$ exceeds it with a given, typically small, probability.

@@

In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM\cite{Toropov1989,Toropov1992,ToropovFilatov1993}) proved to be efficient, e.g. in turbomachinery applications\cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}. This method is an iterative optimization technique based on mid-range approximations built in trust regions. A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated. These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region. The trust region will then translate and change size as optimization progresses. The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values\cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures\cite{ToropovMarkineHolden1999}. The mid-range approximations used in the trust regions, as originally suggested in Ref. \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e. nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure. This was enhanced by the use of gradient-assisted metamodels,\cite{ToropovFilatov1993} use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming\cite{ToropovAlvarez1998}. One of the recent developments\cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e. a two stage approximation building process that is conceptually similar to the original one used in Ref. \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.

The Moving Least Squares Method (MLSM) was proposed in Ref. \cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered data and later used in the mesh-free form of the FEM\cite{Liszka1984}. As suggested in Ref.\cite{ChoiYounYang2001}, it can be used as a technique for metamodelling and used in MDO frameworks. The MLSM is a weighted least squares method where the weights depend on the Euclidian distance from a sample point to where the surrogate model is to be evaluated. The weight value for a certain sample point decays as the distance increases. Describing the weight decay with a Gaussian function tends to be the most useful option even though many others have been evaluated in Ref. \cite{ToropovSchrammSahaiJones2005}. As demonstrated in Ref. \cite{PolynkinToropov2010}, the cross-validated MLSM can be used both for design variable screening and for surrogate modelling. In order to create an efficient MDO framework for problems with disparate discipline attributes Ref. \cite{OllarToropovJones2014} extended the optimization approach of MAM to the use of local DOEs and MLS approximations built in different subspaces of the total design variable space corresponding to the individual disciplines. The subspaces are finally combined into the total design variable space in which the resulting MDO problem is solved.

This paper presents a new development in the Multipoint Approximation Method that makes it capable of handling problems with uncertainty in design variables as well as in additional ‘environmental’ variables. The approach relies on approximations built in the combined space of design variables and environmental variables, and subsequent application of a risk measure and optimization with respect to the deterministic design variables, all within the iterative trust-region-based framework of MAM.

@@

Algorithm 1 (deterministic MAM).
\begin{enumerate}
\item Initialization: choose a starting point $\pmb x^0$ and initial trust region $[\pmb a^0, \pmb b^0]$ such that $\pmb x^0 \in [\pmb a^0, \pmb b^0]$.
\item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k, \pmb b^k] \subset [\pmb a^0, \pmb b^0]$.
  \begin{enumerate}[label=(\alph*)]
    \item Design of Experiments (DoE): a set of points $\pmb x_k^i \in [\pmb a^k, \pmb b^k]$ is chosen to be used for approximation building. Responses are evaluated at the DoE points and approximations are built using the obtained values. Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies\cite{PolynkinToropov2012} and the moving least-squares metamodels\cite{LancasterSalkauskas1981,Liszka1984,ChoiYounYang2001,ToropovSchrammSahaiJones2005}. Other metamodel types could be used as well.

    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x)$ and $\widetilde{F}_j(\pmb x)$, respectively.
    \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
    \begin{equation}
      \label{eq:problem_approx}
      \begin{array}{c}
      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}_0(\pmb x) \\
      s.t.\; \widetilde{F}_j(\pmb x) \le 1,\; j=1,\dots ,M,
      \end{array}
    \end{equation}
    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem   determines the centre of the next trust region.
    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region\cite{KeulenToropovMarkine1996}.
    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $\pmb x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations). If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2.
  \end{enumerate}
  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $\pmb x^{k+1}$.
\end{enumerate}

@@

This choice requires some changes in the MAM algorithm, since metamodels are now built in a larger space than the one where optimisation is performed in. The changes are summarised below.

Algorithm 2 (MAM under uncertainty).
\begin{enumerate}
  \item Initialization: choose a starting point $\pmb x_0 \in {\Bbb R}^n$. Choose an initial trust region $[\pmb a^0,\pmb b^0]$ in the combined space of design variables and environmental variables. Ensure that the initial guess $\pmb x_0$ belongs to the projection of the initial trust region onto the subspace of design variables. The projection of the initial trust region onto the subspace of environmental variables must cover the region where the realisations of the random variable y are likely to be. The coordinates of the trust region corresponding to the environmental variables will not be changed in the course of optimisation. Typically, the trust region along the coordinates corresponding to the $i$-th environmental variable is centred at the mean of the $i$-th environmental variable and has the width of six of its standard deviations.
  \item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k,\pmb b^k]$.
  \begin{enumerate}[label=(\alph*)]
    \item Design of Experiments (DoE): a set of points $(x_i^k,y_i^k)\in [\pmb a^k,\pmb b^k]$ is chosen that will be used to build approximations. Responses are evaluated at the DoE points and approximations are built using the obtained values.
    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x, \pmb y)$ and $\widetilde{F}_j(\pmb x, \pmb y)$, respectively.
    \item The original optimization problem is replaced by the following problem:
    \begin{equation}
      \label{eq:problem_u_tilde}
      \begin{array}{c}
      \min\limits_{a_i \le x_i \le b_i}R_{\pmb u,\pmb y}(\widetilde{F}_0)(\pmb x) \\
      s.t.\; R_{\pmb u,\pmb y}(\widetilde{F}_j)(\pmb x) \le 1,\; j=1,\dots ,M,
      \end{array}
    \end{equation}
    The risk measures are calculated using the metamodels obtained in the previous step. Integrals that arise during this process are calculated with respect to the joint distribution of $\pmb u$ and $\pmb y$. This distribution is assumed to be fixed.
    The approximate problem (\ref{eq:problem_u_tilde}) is solved using Sequential Quadratic Programming (SQP) and the centre of the next trust region $\pmb x^{k+1}$ is determined as the solution of this problem. The optimisation runs in a subspace of the space where metamodels are built: metamodels are defined in the combined space of design variables and environmental variables, whilst optimisation runs in the subspace of the design variables.
    \item The trust region is updated. The size of the trust region along the dimensions corresponding to the environmental variables is unchanged. The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region according to the trust region strategy explained in Ref. \cite{KeulenToropovMarkine1996}.
    \item The termination criterion is checked. If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2
  \end{enumerate}
  \item Optimization terminates. The obtained approximation to the optimum is $\pmb x^{k+1}$.
\end{enumerate}

Summarising, the key changes introduced to account for uncertainties are (i) metamodels built in the combined space of design variables and environmental variables; (ii) optimisation performed in a subspace of the space where metamodels are defined; (iii) additive noise in the design variables; (iv) application of risk measures to transform random responses into deterministic functions.

@@

As it has been already said, a risk measure is a mapping that transforms a random response $F(\pmb x+ \pmb u,\pmb y)$  into a deterministic function that can be optimized. Next we provide some examples of risk measures\cite{RockafellarUryasev2000,SchillingsSchulz2015}.

\begin{itemize}
  \item $R(F)=\mu(F)$ where $\mu(F)$ is the mean of $F$. This choice corresponds to optimising average performance of the system. In typical engineering applications, however, this is not sufficient and more conservative risk measures need to be chosen.
  \item $R(F)=\sup(F)$. This is the most conservative case, in which one is concerned with the worst case performance of the system. The expression $\sup(F)$ stands for the least of all values that are greater than the random response $F$ with probability 1. If the support of the probability density function (PDF) of $F$ is not compact (i.e. $F$ can take arbitrary large values with non-zero probability) than $\sup(F)$ will be infinite.
  \item $R(F)=\mu(F)+k\sigma(F)$, where $\sigma(F)$ is the standard deviation of $F$ and $k>0$ is a constant. This risk measure is sometimes referred to as the safety margin. Optimisation using this risk measure is referred to as robust design optimisation\cite{KochYangGu2004,Yao2011,ZhangZhuChenArendt2013} and reflects the paradigms of a ‘$3\sigma$-design’, a ‘$6\sigma$-design’, etc.
  \item $R(F)=q_F(\alpha)$, where $q_F(\alpha)$ is the value of the quantile function, or simply the quantile, corresponding to a given probability $\alpha\in (0,1)$. This choice corresponds to the paradigm of reliability-based design optimization\cite{AouesChateauneuf2010,Yao2011,YounChoi2004}. In this approach, one wants to keep the probability of the system failing (i.e. violating the constraint or underperforming, depending on the context) below a given threshold. The value $1-\alpha$ is the desired probability of failure. Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset\cite{RockafellarRoyset2010}.
\end{itemize}

For this study, we chose as risk measure the safety margin $R(F)=\mu(F)+k\sigma(F)$, aiming at covering both robust design optimisation and reliability design optimisation. Let us briefly discuss its computation.

@@

The dimension of these integrals is $n+n_e$. To compute them efficiently, methods of high-dimensional integration must be employed. Here, quasi-Monte-Carlo integration\cite{Caflisch1998} based on Niederreiter sequences\cite{Niederreiter1992} is used.

@@

The example considered in this study is a classical engineering optimisation problem known as the scalable cantilevered beam\cite{Vanderpllaats2001}. The engineering object to be optimised is shown in Fig. \ref{fig:beam} taken from Ref. \cite{Vanderpllaats2001}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.45\textwidth]{images/beam.png}
    \caption{The cantilevered beam}
    \label{fig:beam}
\end{figure}

The design variables are the widths $b_i$ and heights $h_i$ of the segments. The number of segments $N$ can be chosen abitrary. The total length of the beam is 500 cm, the lenghths of the segments are $l_i=500/N$  cm. There are $N$ geometric constraints (the aspect ratios of each block (i.e. heights divided by widths) should not exceed 20) and $N$ constraints on the stress, calculated at the left end of each segment (stresses should not exceed  $\bar{\sigma}=14000\; N/cm^2$). There is also also a constraint on the displacement at the tip, which should not exceed 2.5 cm. The load is $P = 50 000\; N$, the Young's modulus is $E=2\cdot 10^7  N/cm^2$.

@@

We are looking for a design with smallest volume $V = \sum_{i=1}^N b_i h_i l_i$. The widths $b_i$ vary from 1.0 to 10.0 cm and the heights $h_i$ from 5.0 to 100.0 cm. The optimisation problem is formulated as follows:
\begin{displaymath}
  \begin{array}{c}
    \min\limits_{\pmb b, \pmb h}V(\pmb b, \pmb h) \\
    s.t.\;1.0\le b_i \le 10.0, \\
    5.0 \le h_i \le 100.0, \\
    y_N\le 2/5, \\
    \sigma_i \le \bar{\sigma}=14000, \\
    \frac{h_i}{b_i}\le 20.
  \end{array}
\end{displaymath}

@@

Recent developments in the Multipoint Approximation Method made it capable of solving large scale problems with uncertainty in parameters modelled as additional (‘environmental’) random variables and uncertainty in the design variables modelled as additive noise. The approach taken combines metamodels built in the combined space of design variables and environmental variables, application of risk measures that map random responses to deterministic functions, and optimisation in a subspace of the space where metamodels are defined, all within the iterative and trust-region-based framework of MAM. Performance is demonstrated on a benchmark example of structural optimisation known as the scalable cantilevered beam.

\section*{Acknowledgement}
The authors are grateful for the support provided by the Russian Science Foundation, project No. 16-11-10150.

@@

\end{document}
__________________________________________________________________________

>>>>

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[misc,geometry]{ifsym}

\usepackage{amsmath}
\usepackage{enumitem}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{epsfig}

@@

\mainmatter

\title{Multipoint Approximation Method for Large Scale Design Optimization Problems Under Uncertainty}
\author{Vassili Toropov \inst{1,3} \and Yury Korolev \inst{2} \and Victor Gergel \inst{3} \and Konstantin Barkalov \inst{3} %\Letter 
\\
\email{konstantin.barkalov@itmm.unn.ru}}

\institute{
Queen Mary University of London, London, UK
\and
University of L\"ubeck, Germany
\and
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
}

@@

\begin{abstract}
The paper presents a new development in the Multipoint Approximation Method (MAM) that makes it capable of handling problems with uncertainty in design variables as well as in additional ‘environmental’ variables. The approach relies on approximations built in the combined space of design variables and environmental variables, and subsequent application of a risk measure and optimization with respect to the deterministic design variables, all within the iterative trust-region-based framework of MAM.

\keywords global optimization, design optimization, multidisciplinary optimization, uncertainty, multipoint approximation method.

\end{abstract}

@@

Optimization under uncertainty deals with responses $F(\pmb x,\pmb y)$ that depend on deterministic design variables $\pmb x$ and random ‘environmental’ variables $\pmb y$. Uncertainty in the design variables can also be modelled by introducing additional random variables, e.g. as additive or multiplicative noise. In this case, the responses $F(\pmb x,\pmb y)$ receive as input the sum $\pmb x+\pmb u$ (if the noise is additive), where $\pmb u$ is a random variable with a known probability distribution, yielding responses depending on $\pmb x,\; \pmb u$ and $\pmb y$. Randomness in $\pmb u$ and $\pmb y$ induces randomness in the responses $F(\pmb x+\pmb u,\pmb y)$. To convert them into deterministic functions, that can be optimized, a mapping $R$, called the \textit{risk measure} \cite{RockafellarUryasev2000} or \textit{robustness measure} \cite{SchillingsSchulz2015}, is applied to the responses. The result is then a deterministic function of the design variables: $\widetilde{F}(\pmb x)=R(F(\pmb x+\pmb u, \pmb y))$. The choice of a particular risk measure $R$ reflects the designer's attitude towards the risk of sub-optimal performance or constraint violation. For example, $R$ can be a combination of the mean and standard deviation of $F(\pmb x+ \pmb u,\pmb y)$ with respect to the joint probability distribution of $\pmb u$ and $\pmb y$ or a quantile of the distribution of $F(\pmb x+ \pmb u,\pmb y)$, i.e. such a value that the random variable of $F(\pmb x+\pmb u,\pmb y)$ exceeds it with a given, typically small, probability.

@@

In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) proved to be efficient, e.g. in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010} . This method is an iterative optimization technique based on mid-range approximations built in trust regions. A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated. These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region. The trust region will then translate and change size as optimization progresses. The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values \cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures \cite{ToropovMarkineHolden1999}. The mid-range approximations used in the trust regions, as originally suggested in \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e. nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure. This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}. One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e. a two stage approximation building process that is conceptually similar to the original one used in \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.

The Moving Least Squares Method (MLSM) was proposed in \cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered data and later used in the mesh-free form of the FEM \cite{Liszka1984}. As suggested in \cite{ChoiYounYang2001}, it can be used as a technique for metamodelling and used in MDO frameworks. The MLSM is a weighted least squares method where the weights depend on the Euclidean distance from a sample point to where the surrogate model is to be evaluated. The weight value for a certain sample point decays as the distance increases. Describing the weight decay with a Gaussian function tends to be the most useful option even though many others have been evaluated in \cite{ToropovSchrammSahaiJones2005}. As demonstrated in \cite{PolynkinToropov2010}, the cross-validated MLSM can be used both for design variable screening and for surrogate modelling. In order to create an efficient MDO framework for problems with disparate discipline attributes \cite{OllarToropovJones2014} extended the optimization approach of MAM to the use of local DOEs and MLS approximations built in different subspaces of the total design variable space corresponding to the individual disciplines. The subspaces are finally combined into the total design variable space in which the resulting MDO problem is solved.

This paper presents a new development in the Multipoint Approximation Method that makes it capable of handling problems with uncertainty in design variables as well as in additional ``environmental'' variables. The approach relies on approximations built in the combined space of design variables and environmental variables, and subsequent application of a risk measure and optimization with respect to the deterministic design variables, all within the iterative trust-region-based framework of MAM.

@@

Algorithm 1 (deterministic MAM).
\begin{enumerate}
\item Initialization: choose a starting point $\pmb x^0$ and initial trust region $[\pmb a^0, \pmb b^0]$ such that $\pmb x^0 \in [\pmb a^0, \pmb b^0]$.
\item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k, \pmb b^k] \subset [\pmb a^0, \pmb b^0]$.
  \begin{enumerate}[label=(\alph*)]
    \item Design of Experiments (DoE): a set of points $\pmb x_k^i \in [\pmb a^k, \pmb b^k]$ is chosen to be used for approximation building. Responses are evaluated at the DoE points and approximations are built using the obtained values. Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies \cite{PolynkinToropov2012} and the moving least-squares metamodels \cite{LancasterSalkauskas1981,Liszka1984,ChoiYounYang2001,ToropovSchrammSahaiJones2005}. Other metamodel types could be used as well.

    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x)$ and $\widetilde{F}_j(\pmb x)$, respectively.
    \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
    \begin{equation}
      \label{eq:problem_approx}
      \begin{array}{c}
      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}_0(\pmb x) \\
      s.t.\; \widetilde{F}_j(\pmb x) \le 1,\; j=1,\dots ,M,
      \end{array}
    \end{equation}
    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem determines the center of the next trust region.
    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $\pmb x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations). If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2.
  \end{enumerate}
  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $\pmb x^{k+1}$.
\end{enumerate}

@@

This choice requires some changes in the MAM algorithm, since metamodels are now built in a larger space than the one where optimization is performed in. The changes are summarized below.

Algorithm 2 (MAM under uncertainty).
\begin{enumerate}
%  \item Initialization: choose a starting point $\pmb x_0 \in {\Bbb R}^n$. 
  \item Initialization: choose a starting point $\pmb x_0 \in {\pmb R}^n$. 
	Choose an initial trust region $[\pmb a^0,\pmb b^0]$ in the combined space of design variables and environmental variables. 
	Ensure that the initial guess $\pmb x_0$ belongs to the projection of the initial trust region onto the subspace of design variables. The projection of the initial trust region onto the subspace of environmental variables must cover the region where the realizations of the random variable y are likely to be. The coordinates of the trust region corresponding to the environmental variables will not be changed in the course of optimization. Typically, the trust region along the coordinates corresponding to the $i$-th environmental variable is centered at the mean of the $i$-th environmental variable and has the width of six of its standard deviations.
  \item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k,\pmb b^k]$.
  \begin{enumerate}[label=(\alph*)]
    \item Design of Experiments (DoE): a set of points $(x_i^k,y_i^k)\in [\pmb a^k,\pmb b^k]$ is chosen that will be used to build approximations. Responses are evaluated at the DoE points and approximations are built using the obtained values.
    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x, \pmb y)$ and $\widetilde{F}_j(\pmb x, \pmb y)$, respectively.
    \item The original optimization problem is replaced by the following problem:
    \begin{equation}
      \label{eq:problem_u_tilde}
      \begin{array}{c}
      \min\limits_{a_i \le x_i \le b_i}R_{\pmb u,\pmb y}(\widetilde{F}_0)(\pmb x) \\
      s.t.\; R_{\pmb u,\pmb y}(\widetilde{F}_j)(\pmb x) \le 1,\; j=1,\dots ,M,
      \end{array}
    \end{equation}
    The risk measures are calculated using the metamodels obtained in the previous step. Integrals that arise during this process are calculated with respect to the joint distribution of $\pmb u$ and $\pmb y$. This distribution is assumed to be fixed.
    The approximate problem (\ref{eq:problem_u_tilde}) is solved using Sequential Quadratic Programming (SQP) and the center of the next trust region $\pmb x^{k+1}$ is determined as the solution of this problem. The optimization runs in a subspace of the space where metamodels are built: metamodels are defined in the combined space of design variables and environmental variables, whilst optimization runs in the subspace of the design variables.
    \item The trust region is updated. The size of the trust region along the dimensions corresponding to the environmental variables is unchanged. The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region according to the trust region strategy explained in \cite{KeulenToropovMarkine1996}.
    \item The termination criterion is checked. If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2
  \end{enumerate}
  \item Optimization terminates. The obtained approximation to the optimum is $\pmb x^{k+1}$.
\end{enumerate}

Summarizing, the key changes introduced to account for uncertainties are (i) metamodels built in the combined space of design variables and environmental variables; (ii) optimization performed in a subspace of the space where metamodels are defined; (iii) additive noise in the design variables; (iv) application of risk measures to transform random responses into deterministic functions.

@@

As it has been already said, a risk measure is a mapping that transforms a random response $F(\pmb x+ \pmb u,\pmb y)$  into a deterministic function that can be optimized. Next we provide some examples of risk measures \cite{RockafellarUryasev2000,SchillingsSchulz2015}.

\begin{itemize}
  \item $R(F)=\mu(F)$ where $\mu(F)$ is the mean of $F$. This choice corresponds to optimizing average performance of the system. In typical engineering applications, however, this is not sufficient and more conservative risk measures need to be chosen.
  \item $R(F)=\sup(F)$. This is the most conservative case, in which one is concerned with the worst case performance of the system. The expression $\sup(F)$ stands for the least of all values that are greater than the random response $F$ with probability 1. If the support of the probability density function (PDF) of $F$ is not compact (i.e. $F$ can take arbitrary large values with non-zero probability) than $\sup(F)$ will be infinite.
  \item $R(F)=\mu(F)+k\sigma(F)$, where $\sigma(F)$ is the standard deviation of $F$ and $k>0$ is a constant. This risk measure is sometimes referred to as the safety margin. Optimization using this risk measure is referred to as robust design optimization \cite{KochYangGu2004,Yao2011,ZhangZhuChenArendt2013} and reflects the paradigms of a ‘$3\sigma$-design’, a ‘$6\sigma$-design’, etc.
  \item $R(F)=q_F(\alpha)$, where $q_F(\alpha)$ is the value of the quantile function, or simply the quantile, corresponding to a given probability $\alpha\in (0,1)$. This choice corresponds to the paradigm of reliability-based design optimization  \cite{AouesChateauneuf2010,Yao2011,YounChoi2004}. In this approach, one wants to keep the probability of the system failing (i.e. violating the constraint or underperforming, depending on the context) below a given threshold. The value $1-\alpha$ is the desired probability of failure. Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset \cite{RockafellarRoyset2010}.
\end{itemize}

For this study, we chose as risk measure the safety margin $R(F)=\mu(F)+k\sigma(F)$, aiming at covering both robust design optimization and reliability design optimization. Let us briefly discuss its computation.

@@

The dimension of these integrals is $n+n_e$. To compute them efficiently, methods of high-dimensional integration must be employed. Here, quasi-Monte-Carlo integration \cite{Caflisch1998} based on Niederreiter sequences \cite{Niederreiter1992} is used.

@@

The example considered in this study is a classical engineering optimization problem known as the scalable cantilevered beam \cite{Vanderpllaats2001}. The engineering object to be optimized is shown in Fig. \ref{fig:beam} taken from \cite{Vanderpllaats2001}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=1.0\textwidth]{beam.png}
    \caption{The cantilevered beam}
    \label{fig:beam}
\end{figure}

The design variables are the widths $b_i$ and heights $h_i$ of the segments. The number of segments $N$ can be chosen arbitrary. The total length of the beam is 500 cm, the lengths of the segments are $l_i=500/N$  cm. There are $N$ geometric constraints (the aspect ratios of each block (i.e. heights divided by widths) should not exceed 20) and $N$ constraints on the stress, calculated at the left end of each segment (stresses should not exceed  $\bar{\sigma}=14000\; N/cm^2$). There is also also a constraint on the displacement at the tip, which should not exceed 2.5 cm. The load is $P = 50 000\; N$, the Young's modulus is $E=2\cdot 10^7  N/cm^2$.

@@

We are looking for a design with smallest volume $V = \sum_{i=1}^N b_i h_i l_i$. The widths $b_i$ vary from 1.0 to 10.0 cm and the heights $h_i$ from 5.0 to 100.0 cm. The optimization problem is formulated as follows:
\begin{displaymath}
  \begin{array}{c}
    \min\limits_{\pmb b, \pmb h}V(\pmb b, \pmb h) \\
    s.t.\;1.0\le b_i \le 10.0, \\
    5.0 \le h_i \le 100.0, \\
    y_N\le 2/5, \\
    \sigma_i \le \bar{\sigma}=14000, \\
    \frac{h_i}{b_i}\le 20.
  \end{array}
\end{displaymath}

@@

Recent developments in the Multipoint Approximation Method made it capable of solving large scale problems with uncertainty in parameters modelled as additional (``environmental'') random variables and uncertainty in the design variables modelled as additive noise. The approach taken combines metamodels built in the combined space of design variables and environmental variables, application of risk measures that map random responses to deterministic functions, and optimization in a subspace of the space where metamodels are defined, all within the iterative and trust-region-based framework of MAM. Performance is demonstrated on a benchmark example of structural optimization known as the scalable cantilevered beam.

\section*{Acknowledgment}
The authors are grateful for the support provided by the Russian Science Foundation, project No. 16-11-10150.

@@

\end{document}
______________________________________________________________________
